Bayesian determination of the effect of a deep eutectic solvent on the structure of lipid monolayers

### 2019-02-19 reports/paper.tex
remove edit comments

The study of lipid monolayers on a non-aqueous solvent like DES is relevant to some potential new scientific areas or technological applications.
For example, it may be possible to study the interactions of proteins with model membranes in a zero or low water environment, or even at low temperatures where aqueous based model membranes are not practical.
To enable such experiments we must first understand how lipid monolayers are affected by a non-aqueous environment.
Published studies covering such non-aqueous systems are, however, rather scarce mirroring the availability of solvents that show a solvophobic effect comparable to water.
In the absence of water, formamide has been reported to support the formation of lipid monolayers \cite{graner_phospholipidic_1995, weinbach_self-assembled_1993}.
Monolayers have also been observed at the air-mercury interface \cite{magnussen_self-assembly_1996, kraack_structure_2002}, although in this case, the similarity to water is somewhat limited.
We have recently shown that self-assembly of surfactant monolayers is possible for DES with behaviour that is much more comparable to that seen in water \cite{arnold_surfactant_2015, sanchez-fernandez_self-assembly_2018} [This paragraph was adapted significantly to reflect the comments of Ref 3].

In this work, we have investigated the structure of phospholipid monolayers at the air-DES interface.
To do this we have used the combination of both X-ray and neutron reflectometry (XRR \& NR), techniques that allow detailed structures to be determined with near-atomic resolution.
In order to interpret our results, we have used a novel analysis method which allows us to sufficiently constrain our model while using Bayesian inference \cite{abeles_sur_1948, parratt_surface_1954} to properly quantify the errors and correlations inherent in such models of reflectometry data.
Specifically, we have constrained the model to ensure that it is chemically self-consistent across multiple measurements at various surface pressures.
Four different phospholipids; 1,2-dipalmitoyl-sn-glycero-3-phosphocholine (DPPC), 1,2-dimyristoyl-sn-glycero-3-phosphocholine (DMPC),  1,2-dilauroyl-sn-glycero-3-phosphocholine (DLPC) and 1,2-dimyristoyl-sn-glycero-3-phospho-(1'-rac-glycerol) (sodium salt) (DMPG), were studied at the interface between a 1:2 mixture of choline chloride:glycerol and air.
This has allowed a comparison of two chemically distinct phospholipid head components, the zwitterionic phosphocholine (PC) and the anionic phosphoglycerol (PG) to be probed in this ionic solvent (Figure \ref{fig:heads}).
In addition, the effect of the tail chain length in the assembly of the Langmuir monolayer was explored by comparing molecules with progressively increasing tail length (12, 14 and 16 carbon atoms in each of the two hydrocarbon chains in each molecule).
Our study is inevitably limited in extent due to the limited availability of X-ray and neutron beamtime together with the prohibitive expense of deuterated chemicals. However, we believe that this work both hints at interesting differences in behaviour for phospholipids monolayers on DES and demonstrates how our approach to fitting can provide a quantitative understanding of this behaviour [This paragraph was added following the comments of Ref 3].
%
\begin{figure}
\includegraphics[width=0.3\textwidth]{figures/head_groups}
\caption{\label{fig:heads}\small The two lipid classes with different head
groups compared in this study, where R indicates the hydrocarbon tail;
(a) phosphatidylglycerol (PG), (b) phosphocholine (PC) [Figure modified to include sodium cation].}
\end{figure}
%

@@

DPPC (C$_{16}$ tails, \SI{>99}{\percent}), DMPC (C$_{14}$ tails, \SI{>99}{\percent}), and the sodium salt of DMPG (C$_{14}$ tails, \SI{>99}{\percent}) [Added salt information for DMPG following Ref 1 comment] were supplied by Avanti Polar Lipids and, DLPC (C$_{12}$ tails, \SI{>99}{\percent}) was supplied by Sigma Aldrich and all were used without further purification.
Deuterated versions of DPPC (d$_{62}$-DPPC, \SI{>99}{\percent}, deuterated tails-only) and DMPC (d$_{54}$-DPPC, \SI{>99}{\percent}, deuterated tails-only) were supplied by Avanti Polar Lipids and used without further purification.
These phospholipids were dissolved in chloroform (\SI{0.5}{\milli\gram\per\milli\liter}) at room temperature.
PC indicates the molcule contains a phosphocholine head component, where PG contains a phosphatidylglycerol head component, these are shown in Figure \ref{fig:heads}.

@@

%
\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/dlpc_ref_sld}
    \includegraphics[width=0.45\textwidth]{figures/dmpc_ref_sld}
    \includegraphics[width=0.45\textwidth]{figures/dppc_ref_sld}
    \includegraphics[width=0.45\textwidth]{figures/dmpg_ref_sld}
    \caption{\small The XRR profiles (left) and SLD profiles (right) for each
  of the four lipids; (a) DLPC, (b) DMPC, (c) DPPC, and (d) DMPG, at the four
  measured surface pressures; see legend above each plot. The different
  surface pressure XRR profiles have been offset in the $y$-axis by an order
  of magnitude and SLD profiles offset in the $y$-axis by
  \SI{5e-6}{\per\square\angstrom}, for clarity.}
    \label{fig:lipids}
\end{figure}
%
\section{Results \& Discussions}
%
The chemically-consistent model was co-refined across the four surface pressure XRR measurements for each lipid.
Although lipid monolayers on water often reach values above \SI{50}{\milli\newton\per\meter}, signifying a decrease in surface tension of ca. \SI{20}{\milli\newton\per\meter} upon addition of lipid, it was not possible to reach those values for the monolayers on DES \cite{mohwald_phospholipid_1990}.
The surface tension of pure choline chloride:glycerol has been previously reported to be \SI{63.5\pm0.5}{\milli\newton\per\meter} \cite{sanchez-fernandez_micellization_2016}.
This means that a surface pressure of ca. \SI{40}{\milli\newton\per\meter} in choline chloride:glycerol will be reached with a reduction of ca. \SI{20}{\milli\newton\per\meter} in the surface tension.
Interestingly, the absolute values of surface tension reduction in both water and choline chloride:glycerol upon addition of lipid are similar.
Therefore, surface pressures of up to \SI{40}{\milli\newton\per\meter} were measured for the monolayers on DES, when possible [This section has been added following comments from Refs 1 \& 3].
The resulting XRR profiles and associated SLD profiles are shown in Figure \ref{fig:lipids}.
Table \ref{tab:liptab} gives details of all varied parameters for each lipid at \SI{30}{\milli\newton\per\meter}, as well as the details of $\phi_h$ which was determined from Eqn. \ref{equ:phih} (the same details for the other surface pressures are available in Section \ref{sec:xrrpara} of the ESI) [Parentheses added following comments from Ref 3].
%
\begin{table*}
    \caption{\label{tab:liptab} The best-fit values, and associated \SI{95}{\percent} confidence intervals for the varying parameters in the XRR models, at the \SI{30}{\milli\newton\per\meter}. The values for $\phi_h$ was obtained from the appropriate use of Eqn. \ref{equ:phih}.}
  \begin{ruledtabular}
    \begin{tabular*}{\textwidth}{ccccc}
    Lipid & DLPC & DMPC & DPPC & DMPG \\
    \hline
    $\sigma$/\si{\angstrom} & \input{../output/dlpc/rough30.txt} & \input{../output/dmpc/rough30.txt} & \input{../output/dppc/rough30.txt} & \input{../output/dmpg/rough30.txt} \\
    $d_t$/\si{\angstrom} & \input{../output/dlpc/tail30.txt} & \input{../output/dmpc/tail30.txt} & \input{../output/dppc/tail30.txt} & \input{../output/dmpg/tail30.txt} \\
    \hline
    $V_t$/\si{\cubic\angstrom} & \input{../output/dlpc/vt.txt} & \input{../output/dmpc/vt.txt} & \input{../output/dppc/vt.txt} & \input{../output/dmpg/vt.txt} \\
    $V_h$/\si{\cubic\angstrom} & \input{../output/dlpc/vh.txt} & \input{../output/dmpc/vh.txt} & \input{../output/dppc/vh.txt} & \input{../output/dmpg/vh.txt} \\
    $d_h$/\si{\angstrom} & \input{../output/dlpc/head.txt} & \input{../output/dmpc/head.txt} & \input{../output/dppc/head.txt} & \input{../output/dmpg/head.txt} \\
    \hline
    $\phi_h$/$\times10^{-2}$ & \input{../output/dlpc/solh30.txt} & \input{../output/dmpc/solh30.txt} & \input{../output/dppc/solh30.txt} & \input{../output/dmpg/solh30.txt} \\
    \end{tabular*}
  \end{ruledtabular}
\end{table*}
%

Following the initial structural determination of the monolayer by XRR, NR was used to confirm the structure and show the applicability of the chemically-consistent model for DPPC and DMPC.
The resulting NR profiles and associated SLD profiles, at a surface pressure of 20 mNm$^{-1}$ are given in Figure \ref{fig:neutron} (the other NR and SLD profiles can be found in Figure \ref{fig:sineutron} of the ESI) [Parentheses added following comments from Ref 3].
Table \ref{tab:neutron} gives details of the varied parameters at each surface pressure as well as $\phi_h$ as determined from NR.
%
\begin{figure}
    \centering
  \includegraphics[width=0.45\textwidth]{figures/dmpc_20n_ref_sld}
    \includegraphics[width=0.45\textwidth]{figures/dppc_20n_ref_sld}
    \caption{\small The NR and SLD profiles at a surface pressure of
  20 mNm$^{-1}$ for two contrasts, the solid line identifing the hdDES contrast and the dashed line the hDES contrast; (a) DMPC, (b) DPPC. The NR profiles have been offset in the $y$-axis by an order of magnitude and SLD profiles offset in the $y$-axis by \SI{5e-6}{\per\square\angstrom}, for clarity.}
    \label{fig:neutron}
\end{figure}
%
%
\begin{table*}
  \caption{\label{tab:neutron} The best-fit values, and associated 95 \%
  confidence intervals for the varying parameters in the co-refined NR models.
  The values of $\phi_h$ were found using Eqn. \ref{equ:phih}.}
  \begin{ruledtabular}
    \begin{tabular*}{\textwidth}{ccccc}
    Lipid & \multicolumn{2}{c}{d$_{54}$-DMPC} & \multicolumn{2}{c}{d$_{62}$-DPPC} \\
    SP/mNm$^{-1}$ & 20 & 25 & 15 & 20 \\
    \hline
    $\sigma_{t,h,s}$/\si{\angstrom} & \input{../output/dmpc/rough20_neutron.txt} & \input{../output/dmpc/rough25_neutron.txt} & \input{../output/dppc/rough15_neutron.txt} & \input{../output/dppc/rough20_neutron.txt} \\
    $d_t$/\si{\angstrom} & \input{../output/dmpc/tail20_neutron.txt} & \input{../output/dmpc/tail25_neutron.txt} & \input{../output/dppc/tail15_neutron.txt} & \input{../output/dppc/tail20_neutron.txt} \\
    \hline
    $\phi_h$/$\times10^{-2}$ & \input{../output/dmpc/solh20_neutron.txt} & \input{../output/dmpc/solh25_neutron.txt} & \input{../output/dppc/solh15_neutron.txt} & \input{../output/dppc/solh20_neutron.txt} \\
    \end{tabular*}
  \end{ruledtabular}
\end{table*}
%
\subsection{Effect of compression on monolayer thickness [Typo removed]}
%
From Tables \ref{tab:liptab} and \ref{tab:neutron}, we can see that, as expected and as found in previous work \cite{mohwald_phospholipid_1990,vaknin_structural_1991}, the thickness of the tail layer increases as the number of carbon atoms in the tail chain increases.
Furthermore, the thickness of the tail layers in these monolayers appears to agree well with values found for water-analogues; \input{../output/dmpc/tail30.txt}\AA\ at 30 mN/m in DES compared with $d_t=15.8$~\si{\angstrom} at \SI{30}{\milli\newton\per\meter} \cite{johnson_structure_1991} in water for DMPC, and \input{../output/dppc/tail30.txt}~\si{\angstrom} at \SI{30}{\milli\newton\per\meter} in DES compared with $d_t=16.7$~\si{\angstrom} at \SI{40}{\milli\newton\per\meter} \cite{helm_phospholipid_1987} in water for DPPC.

The variation of the tail layer thickness in the models with surface pressure is given for each lipid in Figure \ref{fig:lipresults}(e).
As is commonly observed for lipid monolayers at the air-water interface, an increase in surface pressure results in an increase in the thickness of the tail layer.
The variation of this parameter differs for each of the lipids.
For DLPC and DMPC a gradual increase in the thickness is observed at the surface pressures measured here.
For DMPC and DPPC however, after a gradual increase in thickness between \SIlist{15;25}{\milli\newton\per\meter}, a plateau is reached.
This is indicative of the formation of a compacted monolayer where further compression does not produce further changes in the thickness of the tail layer.
This behaviour has been seen for DMPC and DPPC monolayers, where the formation of the condensed phases is reached at high surface concentrations \cite{mohwald_phospholipid_1990} [This description of the plot has been adapted following comments from Ref 1].
%
\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/vh_dt_phih}
    \caption{\small The PDFs of the head volume for each of the four lipids; (a) DLPC, (b) DMPC, (c) DPPC, and (d) DMPG, and the variation of the tail layer thickness (e) and volume fraction of solvent in the head layer (f); DLPC (circles), DMPC (squares), DPPC (triangles), and DMPG (crosses), trend lines have been added between the points to guide the eye. [This figure has been modified slightly to improve clarity].}
    \label{fig:lipresults}
\end{figure}
%

@@

The major difference between the two head groups of the lipids is that the PG is present as a sodium salt, whereas the PC is zwitterionic.
When in solution the anionic PG head is expected to associate with cations in solution, as it does in water \cite{grigoriev_effect_1999} where such interactions depended on a variety of factors including the ionic strength.
In the case of a DES, the environment is inherently ionic and therefore the interaction of an anionic lipid head may be more complex.
As well as interacting with the sodium, the head is likely to interact with the choline cations, similar to behaviour reported previously for surfactant micelles \cite{sanchez-fernandez_counterion_2018}.
The extent of interaction with each of the cations is unclear, but regardless it seems likely that the solvation of the PG head is improved in the DES relative to water.
This better solvation would explain the apparent increase in the volume of the PG head since it would result in a swelling of this group through its strong interactions with the solvent.
In the case of PC, the proximity of a local cation within the molecule results in the same folding of the head group seen in water because this interaction is less transient than the equivalent interactions with the solvent [This paragraph has been adapted significantly following comments from Ref 1].

@@

\subsection{Interparameter correlations}
%
The use of Bayesian inference and MCMC sampling allowed for the probing of the probability distribution function for each parameter individually.
However, it also enables the pairwise inter-parameter PDFs to be investigated, an example for DMPC at \SI{30}{\milli\newton\per\meter} is shown in Figure \ref{fig:dmpcpdfs} (similar plots for the other XRR datasets and the NR datasets are available in Figures \ref{fig:dlpc1} - \ref{fig:dppcn2} of the ESI) [Parentheses added following comments from Ref 3].
These two-dimensional PDFs give important information about the correlations that are present between the parameters of the chemically-consistent model.
The less circular in nature that the 2D PDF is, the greater the correlation that is present between the parameters, e.g. a north-east/south-westerly skewed PDF indicates a positive correlation (where an increase in one parameter correlates an increase in the other) while a north-west/south-easterly skewed PDF indicates a negative correlation (an increase in one correlates with a decrease in the other).
%
\begin{figure}
    \centering
    \includegraphics[width=0.48\textwidth]{figures/dmpc3_all_corner}
    \caption{The multi-parameter PDFs for the chemically-consistent model of DMPC XRR data at \SI{30}{\milli\newton\per\meter}.}
    \label{fig:dmpcpdfs}
\end{figure}
%

>>>>

The study of lipid monolayers on a non-aqueous solvent like DES is relevant to some potential new scientific areas or technological applications.
For example, it may be possible to study the interactions of proteins with model membranes in a zero or low water environment, or even at low temperatures where aqueous based model membranes are not practical.
To enable such experiments we must first understand how lipid monolayers are affected by a non-aqueous environment.
Published studies covering such non-aqueous systems are, however, rather scarce mirroring the availability of solvents that show a solvophobic effect comparable to water.
In the absence of water, formamide has been reported to support the formation of lipid monolayers \cite{graner_phospholipidic_1995, weinbach_self-assembled_1993}.
Monolayers have also been observed at the air-mercury interface \cite{magnussen_self-assembly_1996, kraack_structure_2002}, although in this case, the similarity to water is somewhat limited.
We have recently shown that self-assembly of surfactant monolayers is possible for DES with behaviour that is much more comparable to that seen in water \cite{arnold_surfactant_2015, sanchez-fernandez_self-assembly_2018}.

In this work, we have investigated the structure of phospholipid monolayers at the air-DES interface.
To do this we have used the combination of both X-ray and neutron reflectometry (XRR \& NR), techniques that allow detailed structures to be determined with near-atomic resolution.
In order to interpret our results, we have used a novel analysis method which allows us to sufficiently constrain our model while using Bayesian inference \cite{abeles_sur_1948, parratt_surface_1954} to properly quantify the errors and correlations inherent in such models of reflectometry data.
Specifically, we have constrained the model to ensure that it is chemically self-consistent across multiple measurements at various surface pressures.
Four different phospholipids; 1,2-dipalmitoyl-sn-glycero-3-phosphocholine (DPPC), 1,2-dimyristoyl-sn-glycero-3-phosphocholine (DMPC),  1,2-dilauroyl-sn-glycero-3-phosphocholine (DLPC) and 1,2-dimyristoyl-sn-glycero-3-phospho-(1'-rac-glycerol) (sodium salt) (DMPG), were studied at the interface between a 1:2 mixture of choline chloride:glycerol and air.
This has allowed a comparison of two chemically distinct phospholipid head components, the zwitterionic phosphocholine (PC) and the anionic phosphoglycerol (PG) to be probed in this ionic solvent (Figure \ref{fig:heads}).
In addition, the effect of the tail chain length in the assembly of the Langmuir monolayer was explored by comparing molecules with progressively increasing tail length (12, 14 and 16 carbon atoms in each of the two hydrocarbon chains in each molecule).
Our study is inevitably limited in extent due to the limited availability of X-ray and neutron beamtime together with the prohibitive expense of deuterated chemicals. However, we believe that this work both hints at interesting differences in behaviour for phospholipids monolayers on DES and demonstrates how our approach to fitting can provide a quantitative understanding of this behaviour.
%
\begin{figure}
\includegraphics[width=0.3\textwidth]{figures/head_groups}
\caption{\label{fig:heads}\small The two lipid classes with different head
groups compared in this study, where R indicates the hydrocarbon tail;
(a) phosphatidylglycerol (PG), (b) phosphocholine (PC).}
\end{figure}
%

@@

DPPC (C$_{16}$ tails, \SI{>99}{\percent}), DMPC (C$_{14}$ tails, \SI{>99}{\percent}), and the sodium salt of DMPG (C$_{14}$ tails, \SI{>99}{\percent}) were supplied by Avanti Polar Lipids and, DLPC (C$_{12}$ tails, \SI{>99}{\percent}) was supplied by Sigma Aldrich and all were used without further purification.
Deuterated versions of DPPC (d$_{62}$-DPPC, \SI{>99}{\percent}, deuterated tails-only) and DMPC (d$_{54}$-DPPC, \SI{>99}{\percent}, deuterated tails-only) were supplied by Avanti Polar Lipids and used without further purification.
These phospholipids were dissolved in chloroform (\SI{0.5}{\milli\gram\per\milli\liter}) at room temperature.
PC indicates the molcule contains a phosphocholine head component, where PG contains a phosphatidylglycerol head component, these are shown in Figure \ref{fig:heads}.

@@

%
\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/dlpc_ref_sld}
    \includegraphics[width=0.45\textwidth]{figures/dmpc_ref_sld}
    \includegraphics[width=0.45\textwidth]{figures/dppc_ref_sld}
    \includegraphics[width=0.45\textwidth]{figures/dmpg_ref_sld}
    \caption{\small The XRR profiles (left) and SLD profiles (right) for each
  of the four lipids; (a) DLPC, (b) DMPC, (c) DPPC, and (d) DMPG, at the four
  measured surface pressures; see legend above each plot. The different
  surface pressure XRR profiles have been offset in the $y$-axis by an order
  of magnitude and SLD profiles offset in the $y$-axis by
  \SI{5e-6}{\per\square\angstrom}, for clarity.}
    \label{fig:lipids}
\end{figure}
%
\section{Results \& Discussions}
%
The chemically-consistent model was co-refined across the four surface pressure XRR measurements for each lipid.
Although lipid monolayers on water often reach values above \SI{50}{\milli\newton\per\meter}, signifying a decrease in surface tension of ca. \SI{20}{\milli\newton\per\meter} upon addition of lipid, it was not possible to reach those values for the monolayers on DES \cite{mohwald_phospholipid_1990}.
The surface tension of pure choline chloride:glycerol has been previously reported to be \SI{63.5\pm0.5}{\milli\newton\per\meter} \cite{sanchez-fernandez_micellization_2016}.
This means that a surface pressure of ca. \SI{40}{\milli\newton\per\meter} in choline chloride:glycerol will be reached with a reduction of ca. \SI{20}{\milli\newton\per\meter} in the surface tension.
Interestingly, the absolute values of surface tension reduction in both water and choline chloride:glycerol upon addition of lipid are similar.
Therefore, surface pressures of up to \SI{40}{\milli\newton\per\meter} were measured for the monolayers on DES, when possible.
The resulting XRR profiles and associated SLD profiles are shown in Figure \ref{fig:lipids}.
Table \ref{tab:liptab} gives details of all varied parameters for each lipid at \SI{30}{\milli\newton\per\meter}, as well as the details of $\phi_h$ which was determined from Eqn. \ref{equ:phih} (the same details for the other surface pressures are available in Section \ref{sec:xrrpara} of the ESI).
%
\begin{table*}
    \caption{\label{tab:liptab} The best-fit values, and associated \SI{95}{\percent} confidence intervals for the varying parameters in the XRR models, at the \SI{30}{\milli\newton\per\meter}. The values for $\phi_h$ was obtained from the appropriate use of Eqn. \ref{equ:phih}.}
  \begin{ruledtabular}
    \begin{tabular*}{\textwidth}{ccccc}
    Lipid & DLPC & DMPC & DPPC & DMPG \\
    \hline
    $\sigma$/\si{\angstrom} & \input{../output/dlpc/rough30.txt} & \input{../output/dmpc/rough30.txt} & \input{../output/dppc/rough30.txt} & \input{../output/dmpg/rough30.txt} \\
    $d_t$/\si{\angstrom} & \input{../output/dlpc/tail30.txt} & \input{../output/dmpc/tail30.txt} & \input{../output/dppc/tail30.txt} & \input{../output/dmpg/tail30.txt} \\
    \hline
    $V_t$/\si{\cubic\angstrom} & \input{../output/dlpc/vt.txt} & \input{../output/dmpc/vt.txt} & \input{../output/dppc/vt.txt} & \input{../output/dmpg/vt.txt} \\
    $V_h$/\si{\cubic\angstrom} & \input{../output/dlpc/vh.txt} & \input{../output/dmpc/vh.txt} & \input{../output/dppc/vh.txt} & \input{../output/dmpg/vh.txt} \\
    $d_h$/\si{\angstrom} & \input{../output/dlpc/head.txt} & \input{../output/dmpc/head.txt} & \input{../output/dppc/head.txt} & \input{../output/dmpg/head.txt} \\
    \hline
    $\phi_h$/$\times10^{-2}$ & \input{../output/dlpc/solh30.txt} & \input{../output/dmpc/solh30.txt} & \input{../output/dppc/solh30.txt} & \input{../output/dmpg/solh30.txt} \\
    \end{tabular*}
  \end{ruledtabular}
\end{table*}
%

Following the initial structural determination of the monolayer by XRR, NR was used to confirm the structure and show the applicability of the chemically-consistent model for DPPC and DMPC.
The resulting NR profiles and associated SLD profiles, at a surface pressure of 20 mNm$^{-1}$ are given in Figure \ref{fig:neutron} (the other NR and SLD profiles can be found in Figure \ref{fig:sineutron} of the ESI).
Table \ref{tab:neutron} gives details of the varied parameters at each surface pressure as well as $\phi_h$ as determined from NR.
%
\begin{figure}
    \centering
  \includegraphics[width=0.45\textwidth]{figures/dmpc_20n_ref_sld}
    \includegraphics[width=0.45\textwidth]{figures/dppc_20n_ref_sld}
    \caption{\small The NR and SLD profiles at a surface pressure of
  20 mNm$^{-1}$ for two contrasts, the solid line identifing the hdDES contrast and the dashed line the hDES contrast; (a) DMPC, (b) DPPC. The NR profiles have been offset in the $y$-axis by an order of magnitude and SLD profiles offset in the $y$-axis by \SI{5e-6}{\per\square\angstrom}, for clarity.}
    \label{fig:neutron}
\end{figure}
%
%
\begin{table*}
  \caption{\label{tab:neutron} The best-fit values, and associated 95 \%
  confidence intervals for the varying parameters in the co-refined NR models.
  The values of $\phi_h$ were found using Eqn. \ref{equ:phih}.}
  \begin{ruledtabular}
    \begin{tabular*}{\textwidth}{ccccc}
    Lipid & \multicolumn{2}{c}{d$_{54}$-DMPC} & \multicolumn{2}{c}{d$_{62}$-DPPC} \\
    SP/mNm$^{-1}$ & 20 & 25 & 15 & 20 \\
    \hline
    $\sigma_{t,h,s}$/\si{\angstrom} & \input{../output/dmpc/rough20_neutron.txt} & \input{../output/dmpc/rough25_neutron.txt} & \input{../output/dppc/rough15_neutron.txt} & \input{../output/dppc/rough20_neutron.txt} \\
    $d_t$/\si{\angstrom} & \input{../output/dmpc/tail20_neutron.txt} & \input{../output/dmpc/tail25_neutron.txt} & \input{../output/dppc/tail15_neutron.txt} & \input{../output/dppc/tail20_neutron.txt} \\
    \hline
    $\phi_h$/$\times10^{-2}$ & \input{../output/dmpc/solh20_neutron.txt} & \input{../output/dmpc/solh25_neutron.txt} & \input{../output/dppc/solh15_neutron.txt} & \input{../output/dppc/solh20_neutron.txt} \\
    \end{tabular*}
  \end{ruledtabular}
\end{table*}
%
\subsection{Effect of compression on monolayer thickness}
%
From Tables \ref{tab:liptab} and \ref{tab:neutron}, we can see that, as expected and as found in previous work \cite{mohwald_phospholipid_1990,vaknin_structural_1991}, the thickness of the tail layer increases as the number of carbon atoms in the tail chain increases.
Furthermore, the thickness of the tail layers in these monolayers appears to agree well with values found for water-analogues; \input{../output/dmpc/tail30.txt}\AA\ at 30 mN/m in DES compared with $d_t=15.8$~\si{\angstrom} at \SI{30}{\milli\newton\per\meter} \cite{johnson_structure_1991} in water for DMPC, and \input{../output/dppc/tail30.txt}~\si{\angstrom} at \SI{30}{\milli\newton\per\meter} in DES compared with $d_t=16.7$~\si{\angstrom} at \SI{40}{\milli\newton\per\meter} \cite{helm_phospholipid_1987} in water for DPPC.

The variation of the tail layer thickness in the models with surface pressure is given for each lipid in Figure \ref{fig:lipresults}(e).
As is commonly observed for lipid monolayers at the air-water interface, an increase in surface pressure results in an increase in the thickness of the tail layer.
The variation of this parameter differs for each of the lipids.
For DLPC and DMPC a gradual increase in the thickness is observed at the surface pressures measured here.
For DMPC and DPPC however, after a gradual increase in thickness between \SIlist{15;25}{\milli\newton\per\meter}, a plateau is reached.
This is indicative of the formation of a compacted monolayer where further compression does not produce further changes in the thickness of the tail layer.
This behaviour has been seen for DMPC and DPPC monolayers, where the formation of the condensed phases is reached at high surface concentrations \cite{mohwald_phospholipid_1990}.
%
\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/vh_dt_phih}
    \caption{\small The PDFs of the head volume for each of the four lipids; (a) DLPC, (b) DMPC, (c) DPPC, and (d) DMPG, and the variation of the tail layer thickness (e) and volume fraction of solvent in the head layer (f); DLPC (circles), DMPC (squares), DPPC (triangles), and DMPG (crosses), trend lines have been added between the points to guide the eye.}
    \label{fig:lipresults}
\end{figure}
%

@@

The major difference between the two head groups of the lipids is that the PG is present as a sodium salt, whereas the PC is zwitterionic.
When in solution the anionic PG head is expected to associate with cations in solution, as it does in water \cite{grigoriev_effect_1999} where such interactions depended on a variety of factors including the ionic strength.
In the case of a DES, the environment is inherently ionic and therefore the interaction of an anionic lipid head may be more complex.
As well as interacting with the sodium, the head is likely to interact with the choline cations, similar to behaviour reported previously for surfactant micelles \cite{sanchez-fernandez_counterion_2018}.
The extent of interaction with each of the cations is unclear, but regardless it seems likely that the solvation of the PG head is improved in the DES relative to water.
This better solvation would explain the apparent increase in the volume of the PG head since it would result in a swelling of this group through its strong interactions with the solvent.
In the case of PC, the proximity of a local cation within the molecule results in the same folding of the head group seen in water because this interaction is less transient than the equivalent interactions with the solvent.

@@

\subsection{Interparameter correlations}
%
The use of Bayesian inference and MCMC sampling allowed for the probing of the probability distribution function for each parameter individually.
However, it also enables the pairwise inter-parameter PDFs to be investigated, an example for DMPC at \SI{30}{\milli\newton\per\meter} is shown in Figure \ref{fig:dmpcpdfs} (similar plots for the other XRR datasets and the NR datasets are available in Figures \ref{fig:dlpc1} - \ref{fig:dppcn2} of the ESI).
These two-dimensional PDFs give important information about the correlations that are present between the parameters of the chemically-consistent model.
The less circular in nature that the 2D PDF is, the greater the correlation that is present between the parameters, e.g. a north-east/south-westerly skewed PDF indicates a positive correlation (where an increase in one parameter correlates an increase in the other) while a north-west/south-easterly skewed PDF indicates a negative correlation (an increase in one correlates with a decrease in the other).
%
\begin{figure}
    \centering
    \includegraphics[width=0.48\textwidth]{figures/dmpc3_all_corner}
    \caption{The multi-parameter PDFs for the chemically-consistent model of DMPC XRR data at \SI{30}{\milli\newton\per\meter}.}
    \label{fig:dmpcpdfs}
\end{figure}
%
